---
title: "Loblolly Data Exercises"
author: "Jason Holderieath"
date: "December 11, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Data

`R` includes a dataset that may be more instructive than examples that can be calculated by hand. First, load the packages you expect to need, in this case `library(ggplot2)`. The `Loblolly` dataset can be loaded:

```{r}
library(ggplot2)
data("Loblolly")
Loblolly

```

Looking at this is not particularly instructive. Let's look at the heading.

```{r}
head(Loblolly)
```

From the heading we can identify (from left to right) a column of numbers, most likely some kind of identifier for each _member_. Next is a column named `height`, which appears to be a _quantitative, continuous, variable_. What about `age`? `seed` is a _qualtiative_ variable. How so since it is a number? 

In class we have learned how to sum and square different terms. Let's try some of that.

```{r}
sum(Loblolly$height)
```

What does the sum of height mean? Not much. But the operation could be handy. What about sum of squares? That _will_ be handy.

```{r}
sum(Loblolly$height^2)
```

Is that different than the sum squared?

```{r}
sum(Loblolly$height)^2
```

Sure is. Be careful with that operator and parenthesis placement!

Further we can look at some _descriptive_ statistics.

```{r Loblolly}
summary(Loblolly)
```

We will cover the meaning of these more in the coming weeks. What about graphs? Graphs are cool.

```{r}
# data("Loblolly")
ggplot(Loblolly, aes(y=height, x=age )) + 
  geom_point(aes(colour = Seed))+
  geom_smooth(se = FALSE) +
  labs(title = "Loblolly Height Over Time By Seed Type")
```


## Frequency
We learned about frequency tables. Lets build some more specialized frequency tables with our data.First, lets check our range. Then, starting with height. Say we want to set our breaks every 5 feet. 

```{r}
range(Loblolly$height)
```

Now set the breaks and classify the height according to the half-unit-length sub-intervals with cut. As the intervals are to be closed on the left, and open on the right, we set the right argument as FALSE. 

```{r}
breaks = seq(0, 65, by = 5)
(height.cut = cut(Loblolly$height, breaks, right=FALSE))
```

Compute the frequency of heights in each sub-interval with the table function. 

```{r}
(height.freq = table(height.cut))
```

We apply the cbind function to print the result in column format.
```{r}
cbind(height.freq)
```

```{r}
barplot(height.freq, main = "counts")
```



```{r}
(h.relFreq <- prop.table(height.freq))
```


```{r}
(cumulative.height <- signif(cumsum(h.relFreq), digits = 2))
```

```{r}
barplot(cumulative.height)
```

```{r}
height.ecdf <- ecdf(Loblolly$height)
plot(height.ecdf, xlab = 'Loblolly Tree Height', ylab = '', main = 'Empirical Cumluative Distribution\nLoblolly Tree Height')

```

```{r}
height.ordered = sort(Loblolly$height)
plot(height.ordered, (1:84)/84, type = 's', ylim = c(0, 1), xlab = 'Height', ylab = '', main = 'Empirical Cumluative Distribution\nLoblolly Tree Height')
# mark the 3rd quartile
abline(v = 51.36, h = 0.75)

# add a legend
legend(52, 0.7, '3rd Quartile \n51.36', bty = "n")

# add the label on the y-axis
mtext(text = 'F(x)', side = 2, line = 2.5)


```


```{r}
(tab <- table(Loblolly$Seed))
```

```{r}
barplot(tab, main = "counts")
```

relative frequencies are great. blah blah

```{r}
(relFreq <- signif(prop.table(tab),digits = 2))
```

```{r}
cumsum(relFreq)
```

## Measures of Central Tendency

### Mean
```{r}
mean(Loblolly$height)
#or
sum(Loblolly$height)/nrow(Loblolly)
```

### Median
```{r}
median(Loblolly$height)
#or
n <- (nrow(Loblolly)+1)/2
(height.ordered[n]+height.ordered[n+1])/2
```

### Mode
```{r}
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

Mode(Loblolly$height)
```

## Measures of Dispersion

### Variance
```{r}
var(Loblolly$height, y = NULL, na.rm = FALSE)
```

### Standard Deviation
```{r}
var(Loblolly$height, y = NULL, na.rm = FALSE)^.5
```

## The next big thing

```{r}

```

